---
title: "R Notebook"
output: html_notebook
---
Question 3
Create a table called tab_1 using entry 10 of nodes. Create a table called tab_2 using entry 19 of nodes.

Note that the column names should be c("Team", "Payroll", "Average"). You can see that these column names are actually in the first data row of each table, and that tab_1 has an extra first column No. that should be removed so that the column names for both tables match.

Remove the extra column in tab_1, remove the first row of each dataset, and change the column names for each table to c("Team", "Payroll", "Average"). Use a full_join() by the Team to combine these two tables.

How many rows are in the joined data table?


```{r}
library(rvest)
library(tidyverse)
url <- "https://web.archive.org/web/20181024132313/http://www.stevetheump.com/Payrolls.htm"
h <- read_html(url)
nodes <- html_nodes(h, "table")

tab_1 <- html_table(nodes[10])
tab_1 <- as.data.frame(tab_1)
tab_1 <- tab_1[-1, -1]
tab_2 <- html_table(nodes[19])
tab_2 <- as.data.frame(tab_2)
tab_2 <- tab_2[-1,]
names(tab_1) <- c("Team", "Payroll", "Average")
names(tab_2) <- c("Team", "Payroll", "Average")
full_join(tab_1,tab_2, by = "Team") %>% 
  nrow()


```

Introduction: Questions 4 and 5
The Wikipedia page on opinion polling for the Brexit referendum External link, in which the United Kingdom voted to leave the European Union in June 2016, contains several tables. One table contains the results of all polls regarding the referendum over 2016:
Assign tab to be the html nodes of the "table" class.

How many tables are in this Wikipedia page?

```{r}
library(rvest)
library(tidyverse)
url <- "https://en.wikipedia.org/w/index.php?title=Opinion_polling_for_the_United_Kingdom_European_Union_membership_referendum&oldid=896735054"
target <- read_html(url)
tab <- html_nodes(target,'table')
length(tab)
```

Question 5

Inspect the first several html tables using html_table() with the argument fill=TRUE (you can read about this argument in the documentation). Find the first table that has 9 columns with the first column named "Date(s) conducted".

What is the first table number to have 9 columns where the first column is named "Date(s) conducted"?
```{r}
library(rvest)
library(tidyverse)
url <- "https://en.wikipedia.org/w/index.php?title=Opinion_polling_for_the_United_Kingdom_European_Union_membership_referendum&oldid=896735054"

```

```{r}
# read in raw murders data from Wikipedia
url <- "https://en.wikipedia.org/w/index.php?title=Gun_violence_in_the_United_States_by_state&direction=prev&oldid=810166167"
murders_raw <- read_html(url) %>% 
  html_nodes("table") %>% 
  html_table() %>%
  .[[1]] %>%
  setNames(c("state", "population", "total", "murder_rate"))

# inspect data and column classes
library(dslabs)
data(reported_heights)
class(reported_heights$height)

# convert to numeric, inspect, count NAs
x <- as.numeric(reported_heights$height)
head(x)
sum(is.na(x))

# keep only entries that result in NAs
reported_heights %>% mutate(new_height = as.numeric(height)) %>%
  filter(is.na(new_height)) %>% 
  head(n=10)

# calculate cutoffs that cover 99.999% of human population
alpha <- 1/10^6
qnorm(1-alpha/2, 69.1, 2.9)
qnorm(alpha/2, 63.7, 2.7)

# keep only entries that either result in NAs or are outside the plausible range of heights
not_inches <- function(x, smallest = 50, tallest = 84){
  inches <- suppressWarnings(as.numeric(x))
  ind <- is.na(inches) | inches < smallest | inches > tallest
  ind
}

# number of problematic entries
problems <- reported_heights %>% 
  filter(not_inches(height)) %>%
  .$height
length(problems)

# 10 examples of x'y or x'y" or x'y\"
pattern <- "^\\d\\s*'\\s*\\d{1,2}\\.*\\d*'*\"*$"
str_subset(problems, pattern) %>% head(n=10) %>% cat

# 10 examples of x.y or x,y
pattern <- "^[4-6]\\s*[\\.|,]\\s*([0-9]|10|11)$"
str_subset(problems, pattern) %>% head(n=10) %>% cat

# 10 examples of entries in cm rather than inches
ind <- which(between(suppressWarnings(as.numeric(problems))/2.54, 54, 81) )
ind <- ind[!is.na(ind)]
problems[ind] %>% head(n=10) %>% cat

yes <- as.character(4:7)
no <- as.character(1:3)
s <- c(yes, no)
str_detect(s, "[4-7]")

```

Pattern with Groups
```{r}
# define regex with and without groups
pattern_without_groups <- "^[4-7],\\d*$"
pattern_with_groups <-  "^([4-7]),(\\d*)$"

# create examples
yes <- c("5,9", "5,11", "6,", "6,1")
no <- c("5'9", ",", "2,8", "6.1.1")
s <- c(yes, no)

# demonstrate the effect of groups
str_detect(s, pattern_without_groups)
str_detect(s, pattern_with_groups)

# demonstrate difference between str_match and str_extract
str_match(s, pattern_with_groups)
str_extract(s, pattern_with_groups)

# improve the pattern to recognize more events
pattern_with_groups <-  "^([4-7]),(\\d*)$"
yes <- c("5,9", "5,11", "6,", "6,1")
no <- c("5'9", ",", "2,8", "6.1.1")
s <- c(yes, no)
str_replace(s, pattern_with_groups, "\\1'\\2")

# final pattern
pattern_with_groups <-"^([4-7])\\s*[,\\.\\s+]\\s*(\\d*)$"

# combine stringr commands with the pipe
str_subset(problems, pattern_with_groups) %>% head
str_subset(problems, pattern_with_groups) %>% 
  str_replace(pattern_with_groups, "\\1'\\2") %>% head

```
```{r}
# read raw murders data line by line
filename <- system.file("extdata/murders.csv", package = "dslabs")
lines <- readLines(filename)
lines %>% head()

# split at commas with str_split function, remove row of column names
x <- str_split(lines, ",") 
x %>% head()
col_names <- x[[1]]
x <- x[-1]

# extract first element of each list entry
library(purrr)
map(x, function(y) y[1]) %>% head()
map(x, 1) %>% head()

# extract columns 1-5 as characters, then convert to proper format - NOTE: DIFFERENT FROM VIDEO
dat <- data.frame(parse_guess(map_chr(x, 1)),
                  parse_guess(map_chr(x, 2)),
                  parse_guess(map_chr(x, 3)),
                  parse_guess(map_chr(x, 4)),
                  parse_guess(map_chr(x, 5))) %>%
  setNames(col_names)
  
dat %>% head
  
# more efficient code for the same thing
dat <- x %>%
  transpose() %>%
  map( ~ parse_guess(unlist(.))) %>%
  setNames(col_names) %>% 
  as.data.frame() 

# the simplify argument makes str_split return a matrix instead of a list
x <- str_split(lines, ",", simplify = TRUE) 
col_names <- x[1,]
x <- x[-1,]
x %>% as_tibble() %>%
  setNames(col_names) %>%
  mutate_all(parse_guess)
```

Case Study: Extracting a Table from a PDF
The data come from a paper External link published in the prestigious journal PNAS. However, the data are not provided in a spreadsheet; they are in a table in a PDF document. We could extract the numbers by hand, but this could lead to human error. Instead we can try to wrangle the data using R.

```{r}
library(dslabs)
data("research_funding_rates")
research_funding_rates 

```
Downloading the data
We start by downloading the PDF document then importing it into R using the following code:
```{r}
library("pdftools")
temp_file <- tempfile()
url <- "http://www.pnas.org/content/suppl/2015/09/16/1510159112.DCSupplemental/pnas.201510159SI.pdf"
download.file(url, temp_file)
txt <- pdf_text(temp_file)
file.remove(temp_file)

raw_data_research_funding_rates <- txt[2]

tab <- str_split(raw_data_research_funding_rates, "\n")
tab <- tab[[1]]
the_names_1 <- tab[3]
the_names_2 <- tab[5]
the_names_1 <- the_names_1 %>%
  str_trim() %>%
  str_replace_all(",\\s.", "") %>%
  str_split("\\s{2,}", simplify = TRUE)
the_names_1

 
the_names_2 <-the_names_2 %>%
    str_trim() %>%
    str_split("\\s+", simplify = TRUE)

tmp_names <- str_c(rep(the_names_1, each = 3), the_names_2[-1], sep = "_")
 the_names <- c(the_names_2[1], tmp_names) %>%
     str_to_lower() %>%
     str_replace_all("\\s", "_")
 the_names
 
 new_research_funding_rates <- tab[7:16] %>%
     str_trim %>%
     str_split("\\s{2,}", simplify = TRUE) %>%
     data.frame(stringsAsFactors = FALSE) %>%
     setNames(the_names) %>%
     mutate_at(-1, parse_number)
 new_research_funding_rates %>% head()
```

```{r}
library(rvest)
library(tidyverse)
library(stringr)
url <- "https://en.wikipedia.org/w/index.php?title=Opinion_polling_for_the_United_Kingdom_European_Union_membership_referendum&oldid=896735054"
tab <- read_html(url) %>% html_nodes("table")
polls <- tab[[6]] %>% html_table(fill = TRUE)
polls <- polls[-1,]
names(polls) <- c("dates", "remain", "leave", "undecided", "lead", "samplesize", "pollster", "poll_type", "notes")
length(which(grepl("%", polls$remain)))

#converts the remain vector to a proportion between 0 and 1 from percent?
parse_number(polls$remain)/100
as.numeric(str_replace(polls$remain, "%", ""))/100
```

