---
title: "R Notebook"
output: html_notebook
---

Code: Scatterplot of the relationship between HRs and wins

```{r}
library(Lahman)
library(tidyverse)
library(dslabs)
ds_theme_set()

Teams %>% filter(yearID %in% 1961:2001) %>%
    mutate(HR_per_game = HR / G, R_per_game = R / G) %>%
    ggplot(aes(HR_per_game, R_per_game)) + 
    geom_point(alpha = 0.5)
```

Code: Scatterplot of the relationship between stolen bases and wins
```{r}
Teams %>% filter(yearID %in% 1961:2001) %>%
    mutate(SB_per_game = SB / G, R_per_game = R / G) %>%
    ggplot(aes(SB_per_game, R_per_game)) + 
    geom_point(alpha = 0.5)

```
Code: Scatterplot of the relationship between bases on balls and runs

```{r}
Teams %>% filter(yearID %in% 1961:2001) %>%
    mutate(BB_per_game = BB / G, R_per_game = R / G) %>%
    ggplot(aes(BB_per_game, R_per_game)) + 
    geom_point(alpha = 0.5)
```

Load the Lahman library. Filter the Teams data frame to include years from 1961 to 2001. Make a scatterplot of runs per game versus at bats (AB) per game.
```{r}
Teams %>% filter(yearID %in% 1961:2001)  %>%
    mutate(AB_per_game = AB / G, R_per_game = R / G) %>%
    ggplot(aes(AB_per_game, R_per_game)) + 
    geom_point(alpha = 0.5)

```

Use the filtered Teams data frame from Question 6. Make a scatterplot of win rate (number of wins per game) versus number of fielding errors (E) per game.
```{r}
Teams %>% filter(yearID %in% 1961:2001)  %>%
    mutate(wins_per_game = W / G, err_per_game =  E/ G) %>%
    ggplot(aes(wins_per_game, err_per_game)) + 
    geom_point(alpha = 0.5)
```

Use the filtered Teams data frame from Question 6. Make a scatterplot of triples (X3B) per game versus doubles (X2B) per game.

Which of the following is true?
```{r}
Teams %>% filter(yearID %in% 1961:2001)  %>%
    mutate(triples_per_game = X3B / G, doubles_per_game =  X2B/ G) %>%
    ggplot(aes(triples_per_game, doubles_per_game)) + 
    geom_point(alpha = 0.5)
```
<h3>Correlation</h3>

Load the Lahman library. Filter the Teams data frame to include years from 1961 to 2001.

What is the correlation coefficient between number of runs per game and number of at bats per game?
```{r}
Teams_small <- Teams %>% filter(yearID %in% 1961:2001)
cor(Teams_small$R/Teams_small$G, Teams_small$AB/Teams_small$G)
```

Use the filtered Teams data frame from Question 7.

What is the correlation coefficient between win rate (number of wins per game) and number of errors per game?

```{r}

cor(Teams_small$W/Teams_small$G, Teams_small$E/Teams_small$G)
```

Use the filtered Teams data frame from Question 7.

What is the correlation coefficient between doubles (X2B) per game and triples (X3B) per game?
```{r}
cor(Teams_small$X2B/Teams_small$G, Teams_small$X3B/Teams_small$G)

```

<h3>Anscombe's Quartet/Stratification
</h3>

```{r}
library(tidyverse)
library(HistData)
data("GaltonFamilies")


set.seed(1983)
galton_heights <- GaltonFamilies %>%
  filter(gender == "male") %>%
  group_by(family) %>%
  sample_n(1) %>%
  ungroup() %>%
  select(father, childHeight) %>%
  rename(son = childHeight)

# number of fathers with height 72 or 72.5 inches
sum(galton_heights$father == 72)
sum(galton_heights$father == 72.5)

# predicted height of a son with a 72 inch tall father
conditional_avg <- galton_heights %>%
    filter(round(father) == 72) %>%
    summarize(avg = mean(son)) %>%
    pull(avg)
conditional_avg

# stratify fathers' heights to make a boxplot of son heights
galton_heights %>% mutate(father_strata = factor(round(father))) %>%
    ggplot(aes(father_strata, son)) +
    geom_boxplot() +
    geom_point()

# center of each boxplot
galton_heights %>%
    mutate(father = round(father)) %>%
    group_by(father) %>%
    summarize(son_conditional_avg = mean(son)) %>%
    ggplot(aes(father, son_conditional_avg)) +
    geom_point()

# calculate values to plot regression line on original data
mu_x <- mean(galton_heights$father)
mu_y <- mean(galton_heights$son)
s_x <- sd(galton_heights$father)
s_y <- sd(galton_heights$son)
r <- cor(galton_heights$father, galton_heights$son)
m <- r * s_y/s_x
b <- mu_y - m*mu_x

# add regression line to plot
galton_heights %>%
    ggplot(aes(father, son)) +
    geom_point(alpha = 0.5) +
    geom_abline(intercept = b, slope = m)
```

<h3>Bivariate Normal Distribution</h3>

When a pair of random variables are approximated by the bivariate normal distribution, scatterplots look like ovals. They can be thin (high correlation) or circle-shaped (no correlation).
When two variables follow a bivariate normal distribution, computing the regression line is equivalent to computing conditional expectations.
We can obtain a much more stable estimate of the conditional expectation by finding the regression line and using it to make predictions.

```{r}
galton_heights %>%
  mutate(z_father = round((father - mean(father)) / sd(father))) %>%
  filter(z_father %in% -2:2) %>%
  ggplot() +  
  stat_qq(aes(sample = son)) +
  facet_wrap( ~ z_father)
```

There are two different regression lines depending on whether we are taking the expectation of Y given X or taking the expectation of X given Y.
```{r}
# compute a regression line to predict the son's height from the father's height
mu_x <- mean(galton_heights$father)
mu_y <- mean(galton_heights$son)
s_x <- sd(galton_heights$father)
s_y <- sd(galton_heights$son)
r <- cor(galton_heights$father, galton_heights$son)
m_1 <-  r * s_y / s_x
b_1 <- mu_y - m_1*mu_x

# compute a regression line to predict the father's height from the son's height
m_2 <-  r * s_x / s_y
b_2 <- mu_x - m_2*mu_y
```



```{r}

set.seed(1989) #if you are using R 3.5 or earlier
set.seed(1989, sample.kind="Rounding") #if you are using R 3.6 or later
library(HistData)
library(tidyverse)
library(dslabs)
data("GaltonFamilies")

female_heights <- GaltonFamilies%>%     
    filter(gender == "female") %>%     
    group_by(family) %>%     
    sample_n(1) %>%     
    ungroup() %>%     
    select(mother, childHeight) %>%     
    rename(daughter = childHeight)

#Daughter's height from mother’s height.
mu_m <- mean(female_heights$mother)
mu_d <- mean(female_heights$daughter)
s_m <- sd(female_heights$mother)
s_d <- sd(female_heights$daughter)

#Correlation coefficient
r <- cor(female_heights$mother, female_heights$daughter)

#Slope of regression line predicting daughters' height from mothers' heights
m_1 <-  r * s_d / s_m

#Intercept of regression line predicting daughters' height from mothers' heights
b_1 <- mu_d - m_1*mu_m

#Change in daughter's height in inches given a 1 inch increase in the mother's height
r * s_y/s_x

#What percent of the variability in daughter heights is explained by the mother's height?
r^2 * 100

```

A mother has a height of 60 inches.

Using the regression formula, what is the conditional expected value of her daughter's height given the mother's height?

```{r}
 m = r * s_y/s_x
b = mu_y - (r * s_y/s_x)*mu_x
x = 60
m*x+b

```

